---
title: "STAT 9210 Final Project"
author: "Pragyat (Prag) Agrawal and Sahil Kulkarni"
date: "2023-04-25"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r Setup, include=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(echo = T, fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output

# Package setup
if(!require("pacman")) install.packages("pacman")

#install.packages("fGarch", repos = "http://cran.us.r-project.org")
#libary("fGarch")

pacman::p_load(plyr, tidyverse, dplyr, ggplot2, ggthemes, data.table, lubridate,
               GGally, RColorBrewer, ggsci, plotROC, usmap,
               plotly, ggpubr, vistime, pander, softImpute, sjmisc, sjPlot, stringr,
               hwwntest, lmtest, spelling, tseries, moments, fGarch, exactRankTests, FSA, psych,
               lme4, lmerTest, nlme, car, readxl)

options(tinytex.verbose = TRUE)


```


# Introduction

Breathing practices are vital for humans as they influence how we feel and shape human behavior. But scientific evidence on breathing and its effect on mind-body connection have been limited. Structured breathing practices have their own share of benefits as they help with enhancing mood and reduce physiological arousal in humans. There have been several studies which have established this. We seek to study this same effect and replicate its result by applying the concepts we learned in class to find if long breathing exercises can help people reach mindfulness and reduce stress and anxiety. 

# Study

We looked into a study by Cell Reports Medicine which according to their website cell.com is "premium open-access journal from Cell Press publishing cutting-edge research in translational and clinical biomedical sciences that inform and influence human health and medicine."


```{r, include=FALSE}
filename = read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

view(filename)
head(filename)
```

## Participants

Throughout the study, 140 potential targets were consented but 114 were invited to participate in the study. There was some attrition because of pandemic related reasons or loss of contact with the target. According to the study, The general ease of following instructions and performing the interventions and subjective experience of the interventions were assessed by an optional debriefing questionnaire at the end of the study. 

## Treatment

The experiment put the 114 participants to a breathing and meditation protocol and measured the affect of that on their state anxiety. Specifically, it measured positive affect (Positive And Negative Affect Schedule (PANAS)) and state anxiety (State-Trait Anxiety Inventory (STAI)) scores on each participant before and after the experiment. Mindfulness meditation and breathwork groups both experienced an increase in daily positive affect.


## Data Visualization

```{r}
attach(filename)
filename$`Sleep Score` <- as.numeric(filename$`Sleep Score`)
filename$`Hours of Sleep` <- as.numeric(filename$`Hours of Sleep`)
# create new variables with different names
my_hours <- filename$`Hours of Sleep`
my_score <- filename$`Sleep Score`

# use the original variables with the `::` operator
ggplot(filename, aes(x = my_hours, y = my_score)) + 
  geom_point() + 
  geom_smooth(method = lm, color = 'red', se = F) + 
  ylim(NA, 100) + 
  annotate("text",x=10,y=10,label=(paste0("slope==",coef(lm(filename$`Sleep Score`~filename$`Hours of Sleep`))[2])),parse=TRUE)
```
This graphs shows that sleep score is positively related to the number of hours a person sleeps. This means a higher number of hours sleeping, corresponds to a higher sleep score. This makes sense as the quality of sleep increases with the quantity of sleep a person gets. 


```{r}
colnames(filename)

```

```{r, include = FALSE}
library(ggplot2)
library(tidyr)
library(cowplot)

variables <- filename[,4:9]
variables <- apply(variables, 2, function(x) as.numeric(as.character(x)))
```

```{r}



variables <- variables[!apply(is.na(variables) | variables == "", 1, all),]
variables_df <- as.data.frame(variables)

group1 <- variables_df[, c(1, 2, 5)]
group2 <- variables_df[, c(3, 4, 6)]

group1_melted <- gather(group1, key = "variable", value = "value")
group2_melted <- gather(group2, key = "variable", value = "value")

p1 <- ggplot(group1_melted, aes(x = variable, y = value)) +
  geom_boxplot(position = position_dodge(0.75)) +
  xlab("") +
  ylab("Group 1") + 
  theme(axis.text.x = element_text(angle = 10))


p2 <- ggplot(group2_melted, aes(x = variable, y = value)) +
  geom_boxplot(position = position_dodge(0.75)) +
  xlab("") +
  ylab("Group 2")+ 
  theme(axis.text.x = element_text(angle = 10))

plot_grid(p1, p2, ncol = 2)

```
Looking at the results, we see the breathwork group had a notably higher increase in daily positive affect, a reduction in daily negative affect and a reduction in the state anxiety. All of these indicate that there is some positive effect because of the breathing and meditation protocol that has been assigned to the participants. This is proven by the regression below which shows Mindful Meditation as a significant indicator for a reduction in anxiety and an increase in positive effect. 

## Mixed-effect Model

Here we use Linear mixed-effect to show the positive effect change based upon the type of exercise, and number of days over which the exercise was performed. Each participant id is treated as a random effect to account for inter-participant variability over the number of days. 

Linear mixed-effect models are a type of regression model that are used when there are both fixed (exercise type) and random(participant id) effects in the data. The mixed-effect model provides estimates of the fixed effects and the variance components of the random effects, as well as goodness-of-fit measures and other diagnostic statistics. The lmerTest package was used to obtain p-values and confidence intervals for the fixed effects.

We found that Days from Round 1 (i.e. increase in number of days following a practice) helps in increase the positive effect on the participant (indicator for goodness/wellness).

```{r}
library(readxl)
library(lme4)

library(lmerTest)
library(tidyverse)
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

filename[filename=='.'] <- NA

data_set <- filename[, c(1, 2, 3, 4, 6)]

data_set <- data_set %>% drop_na()

data_set$`Post PANAS Positive Affect` <- as.numeric(data_set$`Post PANAS Positive Affect`)
data_set$`Pre PANAS Positive Affect` <- as.numeric(data_set$`Pre PANAS Positive Affect`)
data_set$`Round 1 Exercise` <- as.factor(data_set$`Round 1 Exercise`)
data_set$`Days from  Round1 Day1` <- as.numeric(data_set$`Days from  Round1 Day1`)


data_set$change <- data_set$`Post PANAS Positive Affect` - data_set$`Pre PANAS Positive Affect`

model = lmer(change ~ `Days from  Round1 Day1` + `Round 1 Exercise` * `Days from  Round1 Day1` +
               `Round 1 Exercise` + (1|ScrSubjectID),
             data=data_set,
             REML=TRUE)
summary(model)

anova(model)
```

```{r Trait_z_1, include=FALSE}
#Trait anxiety score analysis
library('ggpubr')
library(readxl)
trait <- data.frame(matrix(ncol =5, nrow = 120 ))


colnames(trait) <- c("subjectID", "Pre Trait", "Post Trait", "Group", "Type")

# filename = BWPilot_CombinedData_20210308
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")


for (i in 1:length(unique(filename$`ScrSubjectID`))){
  print(i)
  id <- unique(filename$`ScrSubjectID`)[i]
  print (id)
  b <- filename$`ScrSubjectID` == as.numeric(id) 
  a <- filename[b,] # data for each individuaal subject
  
  
  round1pre<- subset(a, `Days from  Round1 Day1`<1)

  round1post<- subset(a, `Days from  Round1 Day1`> 1 &`Days from  Round1 Day1`< 31 )
  prepromisind <-round1pre$`Full STAI Trait Anxiety` != "."
  prepromis <- as.numeric(round1pre$`Full STAI Trait Anxiety`[prepromisind])
  postpromisind <-round1post$`Full STAI Trait Anxiety` != "."
  postpromis <- as.numeric(round1post$`Full STAI Trait Anxiety`[postpromisind])
  trait$subjectID [i] <- id
  if (length(prepromis)>0) {
    trait$`Pre Trait` [i] <- prepromis[length(prepromis)]
    
  }
  
  if (length(postpromis)>0) {
    trait$`Post Trait` [i] <- postpromis[length(postpromis)]
    
  }
  trait$Group [i] <-  a$`Round 1 Exercise`[1]
  if (trait$Group[i] == 'Mindful Meditation'){trait$Type[i]<- 'Mindfulness'}
    else {trait$Type[i]<- 'Breathwork'}
  
}
```

```{r Trait_z_2}

effect_trait <- as.numeric(trait$`Post Trait`) - as.numeric(trait$`Pre Trait`)

trait['delta trait']<- effect_trait


b <- trait$Group == "Mindful Meditation"
trait_medi <- trait[b,]

b <- trait$Group == "Box Breathing"
trait_box <- trait[b,]

b <- trait$Group == "Super Oxygenation"
trait_superox <- trait[b,]

b <- trait$Group == "Slow Breathing"
trait_slow <- trait[b,]
####

effect_superox <- as.numeric(trait_superox$`Post Promis`)- as.numeric(trait_superox$`Pre Promis`)
effect_medi <- as.numeric(trait_medi$`Post Promis`)- as.numeric(trait_medi$`Pre Promis`)
effect_box <- as.numeric(trait_box$`Post Promis`)- as.numeric(trait_box$`Pre Promis`)
effect_slow<- as.numeric(trait_slow$`Post Promis`) - as.numeric(trait_slow$`Pre Promis`)


wilcox.test(trait_superox$`Post Trait`,trait_superox$`Pre Trait`, paired = TRUE)
wilcox.test(trait_slow$`Post Trait`,trait_slow$`Pre Trait`, paired = TRUE)
wilcox.test(trait_box$`Post Trait`,trait_box$`Pre Trait`, paired = TRUE)
wilcox.test(trait_medi$`Post Trait`,trait_medi$`Pre Trait`, paired = TRUE)

kruskal.test(`delta trait`~Group, data=trait)
wilcox.test(trait_slow$`Post Trait`,trait_slow$`Pre Trait`, paired = TRUE) #WILCOX IS 2 SAMPLE ONLY, KRUSKAL WALLIS MORE) , ~


#round 1
trait <- trait [complete.cases(trait),]
trait$Group<- factor(trait$Group, levels = c('Mindful Meditation', 'Slow Breathing', 'Box Breathing', 'Super Oxygenation'))
e <- ggplot(trait, aes(x = Group, y = `delta trait`))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = "grey20"), size = 2, position = position_jitterdodge(),show.legend = F) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())

## by type t test

traitmed<- subset(trait, Type =='Mindfulness')
traitbreath <- subset(trait, Type == 'Breathwork')
t.test(traitmed$`delta trait`, y =traitbreath$`delta trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)


t.test(traitmed$`Pre Trait`, y =traitmed$`Post Trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.95)

t.test(traitbreath$`Pre Trait`, y =traitbreath$`Post Trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.95)

traitmed<- traitmed[complete.cases(traitmed),]
mean (traitmed$`delta trait`)/sd(traitmed$`Pre Trait`)
traitbreath<- traitbreath[complete.cases(traitbreath),]
mean (traitbreath$`delta trait`)/sd(traitbreath$`Pre Trait`)

wilcox.test(traitbreath$`Post Trait`,traitbreath$`Pre Trait`, paired = TRUE)
wilcox.test(traitbreath$`delta trait`,traitmed$`delta trait`, paired = FALSE)

plot.new() 
for (i in 1:nrow(traitbreath)){
  if (is.na(traitbreath$`Pre Trait`[i])|is.na(traitbreath$`Post Trait`[i])) next
  par(new=T)
  plot(c(traitbreath$`Pre Trait`[i],traitbreath$`Post Trait` [i]), type= 'o', ylab = 'Trait Anxiety', xlab = 'All Breathwork', xlim=c(0,3),ylim=c(20,80), xaxt='n')
}

axis(1, at=c(1,2), labels=c('Before', 'After'))

trait$Type<- factor(trait$Type, levels = c('Mindfulness', 'Breathwork'))
trait<- trait[complete.cases(trait),]
e <- ggplot(trait, aes(x = Type, y = `delta trait`))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = "grey20"), size = 2, position = position_jitterdodge(),show.legend = F) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())

kruskal.test(`delta trait`~Type, data = trait)

#ttest by group

traitbreath <- subset(trait, Group =='Box Breathing')
traitmed<- subset(trait, Group =='Super Oxygenation')


t.test(traitmed$`delta trait`, y =traitbreath$`delta trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```
We have conducted Wilcoxin and Kruskal-Wallis rank sum test, whose output is above. The wilcoxin test for the super-oxygenation group is significant at the 0.10 level. The Kruskal-Wallis test is not significant which may indicate that there is no evidence that the medians of the pre and post groups are significantly different. However, it does not say that there are no differences in the groups. It may be the case that the differences are too small to be detected.

The paired t-test for pre and post traitbreath shows that the difference in factor is significant. 

This graphs shows the before and after trait anxiety. This graph does show  clearly that there is a reduction in before and after state anxiety.


```{r SlopesRR_1, include=FALSE}
library("ggpubr")
library(readxl)

filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

slopes<- data.frame(matrix (ncol = 11, nrow = 120))

for (i in 1:length(unique(filename$`ScrSubjectID`))){
  print(i)
  id <- unique(filename$`ScrSubjectID`)[i]
  print (id)
  b <- filename$`ScrSubjectID` == as.numeric(id) 
  a <- filename[b,] # data for each individuaal subject
  c<- a #round 1
  
  c[c=='.'] <- NA
  
  if (nrow(c) == 0){
    next()
  }
  
  if (sum(is.na(c$RHR))==length(c$RHR))
    next()
  
  #  if (sum(!is.na(c$RHR))<2) 
   #   next()
   
  if (sum(!is.na(c$`Pre PANAS Positive Affect`))<2) 
      next()
  
  z<- c[,c(2,4,5,6,7,8,9,15,16,20)] # extract hrv, rhr, respiratory rate, sleep scores
  z<- unique.data.frame(z)

  #reg1 <- lm(c$HRV ~ c$`Days from  Round1 Day1`,data=c, na.action = na.omit)
  slope <- lm(formula = z$HRV ~ z$`Days from  Round1 Day1`)
  slopeHRV<-slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  slope <- lm(formula = z$RHR ~ z$`Days from  Round1 Day1`)
  slopeRHR<- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  slope <- lm(formula = z$`Respiration Rate` ~ z$`Days from  Round1 Day1`)
  slopeRR <- slope[['coefficients']][['z$`Days from  Round1 Day1`']]

 slope <- lm(formula = c$`Pre PANAS Positive Affect`~ z$`Days from  Round1 Day1`)
slopeprepanaspos<- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  
  
  z$deltapanaspos<- as.numeric(z$`Post PANAS Positive Affect`)- as.numeric(z$`Pre PANAS Positive Affect`)
    slope <- lm(formula = z$deltapanaspos~ as.numeric(z$`Days from  Round1 Day1`))
    slopedeltapanaspos <- slope[['coefficients']][2]

  slopes [i,1] <- id
  slopes [i,2] <- slopeHRV
  slopes [i,3] <- slopeRHR
  slopes [i,4] <- slopeRR
 slopes [i,5] <- slopedeltapanaspos
  slopes [i,6] <- c$`Round 1 Exercise`[1]

  slopes [i,11]<- slopeprepanaspos
  
  colnames(slopes)<- c("subjectID", "slopeHRV", "slopeRHR", "slopeRR","slopedeltapanaspos","Group", "SleepScore", "SleepEfficiency", "HoursOfSleep", "slopeprestate", "slopeprepanaspos")
  
}

```

```{r SlopesRR_2}

#remove any empty columns

slopes <- slopes[,colSums(is.na(slopes))<nrow(slopes)]
slopes_rhrhrvRR_excludinglessthan8datapoints<- slopes # we are not excluding data. 

#statistical tests
#non-parametric

slopes_rhrhrvRR_excludinglessthan8datapoints<- slopes_rhrhrvRR_excludinglessthan8datapoints[complete.cases(slopes_rhrhrvRR_excludinglessthan8datapoints),]
kruskal.test(slopeRR~Group, data = slopes_rhrhrvRR_excludinglessthan8datapoints)
pairwise.wilcox.test(slopes_rhrhrvRR_excludinglessthan8datapoints$slopeRR, slopes_rhrhrvRR_excludinglessthan8datapoints$Group, paired = FALSE)

    
  for (i in 1:nrow(slopes_rhrhrvRR_excludinglessthan8datapoints)){
  if (slopes_rhrhrvRR_excludinglessthan8datapoints$Group[i] =="Mindful Meditation"){ slopes_rhrhrvRR_excludinglessthan8datapoints$Type[i]<- 'Mindfulness'}
  # else if (c$`Round 1 Exercise` == 'Box Breathing'){ weekly$Type[i]<- 'Box Breathing'}
  else {slopes_rhrhrvRR_excludinglessthan8datapoints$Type[i]<- 'Breathing'}
  }
  slopes_rhrhrvRR_excludinglessthan8datapoints$Type<- factor(slopes_rhrhrvRR_excludinglessthan8datapoints$Type, levels = c("Mindfulness", "Breathing"))
  
  # kruskal.test(slopeRR~Type, data = slopes_rhrhrvRR_excludinglessthan8datapoints)
  
  rrm <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Mindfulness")
  rrb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Breathing")
  rrsb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Group == "Slow Breathing")
  rrbb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Group == "Box Breathing")
  rrso <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Group == "Super Oxygenation")
  
  deltapanasposm <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Mindfulness")
  deltapanasposb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Breathing")
  
# t test
  
  t.test(rrm$slopeRR, y =rrb$slopeRR,
         alternative = c ("two.sided", "less", "greater"),
         mu = 0, paired = FALSE, var.equal = FALSE,
         conf.level = 0.95)
  signal<- slopes_rhrhrvRR_excludinglessthan8datapoints
  signal <- signal [complete.cases(signal),]
  signal$Type<- factor(signal$Type, levels = c('Mindfulness', 'Breathing'))
  signal$Group<- factor(signal$Group, levels = c('Mindful Meditation', 'Slow Breathing', 'Box Breathing', 'Super Oxygenation'))
  print(i)

  e <- ggplot(signal, aes(x = Group, y = slopeRR))+geom_boxplot(alpha = 0.3)
  e
 
```
Here we have boxplots behind the mindulness and breathing exercises. The boxplot is slopeRR vs the different kinds of breaathing exercises. It shows that there are four groups and two of the groups (except box breathing and super oxygenation) show that there is a positive effect between them and the slopeRR. SlopeRR corresponds to the effect of the number of days we stick to a regime. Non-zero slopes show that all exercises are beneficial in long run (corresponding to how long we stick with a regime). 


## Matching

Here we consider matching participants from a group where group is defined by type of exercise performed by the participants to, a control group the control group is defined by all the participants who enrolled in the study. The control group is chosen as such because we have data for all the participants before they started doing any of the recommended exercises. We can match each treatment group to all the participants. The control group and the treatment group will be matched based on: RHR, HRV, Sleep Score, Hours of sleep, sleep efficacy, hours of sleep and respiration rate Outcomes before and after the exercises will be chosen for determining treatment effect

```{r}

#-------------------------------------------------------------------------------
#Preparing dataset
library(tidyverse)
library(readxl)
library(dplyr)
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")
filename[filename=='.'] <- NA
data_set <- filename[, c(1,2,15,16,17,18,19,20)]


data_set <- data_set %>% drop_na()

data_set$`Days from  Round1 Day1` <- as.numeric(data_set$`Days from  Round1 Day1`)

data_set <- data_set[-20 < data_set$`Days from  Round1 Day1`, ]

data_set$RHR <- as.numeric(data_set$RHR)
data_set$HRV <- as.numeric(data_set$HRV)
data_set$`Sleep Score` <- as.numeric(data_set$`Sleep Score`)
data_set$`Hours of Sleep` <- as.numeric(data_set$`Hours of Sleep`)
data_set$`Sleep Efficiency` <- as.numeric(data_set$`Sleep Efficiency`)
data_set$`Respiration Rate` <- as.numeric(data_set$`Respiration Rate`)

t.first <- filename[match(unique(data_set$ScrSubjectID), data_set$ScrSubjectID),c(3)]
t.sec <- ifelse(t.first$`Round 1 Exercise` == "Box Breathing", 1,
                ifelse(t.first$`Round 1 Exercise` == "Slow Breathing", 2,
                       ifelse(t.first$`Round 1 Exercise` == "Super Oxygenation", 3, 4)) 
                )

a <- rep(0, 113)

treatment <- c(rbind(a, t.sec))

treatment <- treatment[-122]
treatment <- treatment[-137]
data_set <- data_set %>% 
  mutate(pos = `Days from  Round1 Day1`>0)

data_set_matching <- data_set %>% group_by(ScrSubjectID, pos) %>%
  summarise_all(mean)

data_set_matching <- data.frame(data_set_matching, treatment)
#-------------------------------------------------------------------------------
library(tidyverse)
library(ggplot2)
library(MASS)
#-------------------------------------------------------------------------------
smahal=
  function(z,X){
    X<-as.matrix(X)
    n<-dim(X)[1]
    rownames(X)<-1:n
    k<-dim(X)[2]
    m<-sum(z)
    for (j in 1:k) X[,j]<-rank(X[,j])
    cv<-cov(X)
    vuntied<-var(1:n)
    rat<-sqrt(vuntied/diag(cv))
    cv<-diag(rat)%*%cv%*%diag(rat)
    out<-matrix(NA,m,n-m)
    Xc<-X[z==0,]
    Xt<-X[z==1,]
    9
    rownames(out)<-rownames(X)[z==1]
    colnames(out)<-rownames(X)[z==0]
    icov<-ginv(cv)
    for (i in 1:m) out[i,]<-mahalanobis(Xc,Xt[i,],icov,inverted=T)
    out
  }

addcaliper=function(dmat,z,logitp,calipersd=.5,penalty=1000){
  # Pooled within group standard devation
  sd.logitp=sqrt((sd(logitp[z==1])^2+sd(logitp[z==0])^2)/2)
  adif=abs(outer(logitp[z==1],logitp[z==0],"-"))
  adif=(adif-(calipersd*sd.logitp))*(adif>(calipersd*sd.logitp))
  dmat=dmat+adif*penalty
  dmat
}
#-------------------------------------------------------------------------------
subset <- data_set_matching[data_set_matching$treatment %in% c(0,1), ]  
datatemp=subset

subset$Hours.of.Sleep
subset$Sleep.Efficiency
subset$Respiration.Rate

subset$Days.from..Round1.Day1

colnames(subset)[3] <- "days"

#Score model
propscore.model=glm(pos ~ RHR + HRV  +Sleep.Score+
                     Respiration.Rate,
                    family=binomial,x=TRUE,y=TRUE,
                    data=subset)
# #-----------------------------------------------------------------------------
datatemp$outcome = datatemp$Sleep.Efficiency 
datatemp$treated=propscore.model$y
datatemp$treatment=datatemp$treated

# Use the caret package to include all categories of categorical variables (i.e.,
# do not leave out one category) in X matrix
library(caret)
dmy=dummyVars(propscore.model$formula,data=datatemp)
Xmat=data.frame(predict(dmy,newdata=datatemp))
# Matrix of covariates to include in the Mahalanobis distance, based on prognostically
#important variables
Xmatmahal=Xmat
treated=datatemp$treated
datatemp$logit.ps=predict(propscore.model)
# Use Hansen (2009)’s rule for removing subjects who lack overlap
# from Notes 6
logit.propscore=predict(propscore.model)
pooled.sd.logit.propscore=sqrt(var(logit.propscore[datatemp$treatment==1])/2+var(logit.propscore[datatemp$treatment==0])/2)
min.treated.logit.propscore=min(logit.propscore[datatemp$treatment==1])
max.control.logit.propscore=max(logit.propscore[datatemp$treatment==0])
# How many treated and control subjects lack overlap by Hansen's criterion
no.treated.lack.overlap=sum(logit.propscore[datatemp$treatment==1]>(max.control.logit.propscore+.5*pooled.sd.logit.propscore))
no.control.lack.overlap=sum(logit.propscore[datatemp$treatment==0]<(min.treated.logit.propscore-.5*pooled.sd.logit.propscore))
# If there are subjects who lack overlap, remove them from the datatemp dataset
datatemp.original=datatemp
datatemp.full=datatemp
Xmat.original=Xmat
Xmat.full=Xmat
if(no.treated.lack.overlap+no.control.lack.overlap>0){
  which.remove=which((logit.propscore>(max.control.logit.propscore+.5*pooled.sd.logit.propscore))|(logit.propscore<(min.treated.logit.propscore-.5*pooled.sd.logit.propscore)))
  datatemp=datatemp[-which.remove,]
  datatemp.full=rbind(datatemp,datatemp.original[which.remove,])
  Xmat=Xmat[-which.remove,]
  Xmat.full=rbind(Xmat,Xmat.original[which.remove,])
  Xmatmahal=Xmatmahal[-which.remove,]
}
rownames(datatemp)=seq(1,nrow(datatemp),1)

# Rank based Mahalanobis distance
distmat=smahal(datatemp$treated,Xmatmahal)
# Add caliper
distmat=addcaliper(distmat,datatemp$treated,datatemp$logit.ps,calipersd=0.5)
# Label the rows and columns of the distance matrix by the rownames in datatemp
rownames(distmat)=rownames(datatemp)[datatemp$treated==1]
colnames(distmat)=rownames(datatemp)[datatemp$treated==0]
#-----------------#-----------------#-----------------#-----------------
library(optmatch)
matchvec=fullmatch(distmat,data=datatemp) 
datatemp$matchvec=matchvec 
# Number the strata
matchedset.index=substr(matchvec,start=3,stop=10)
matchedset.index.numeric=as.numeric(matchedset.index)
# Have matchedset.index.numeric.full append 0 to matchedset index.numeric for
# the removed subjects
if(no.control.lack.overlap+no.treated.lack.overlap==0){
  matchedset.index.numeric.full=matchedset.index.numeric
}
if(no.control.lack.overlap+no.treated.lack.overlap>0){
  matchedset.index.numeric.full=c(matchedset.index.numeric,rep(0,no.control.lack.overlap+no.treated.lack.overlap))
}

standardized.diff.harmonic.func=function(x,treatment,stratum.myindex,missing=rep(0,length(x))){
  xtreated=x[treatment==1 & missing==0];
  xcontrol=x[treatment==0 & missing==0];
  var.xtreated=var(xtreated);
  var.xcontrol=var(xcontrol);
  combinedsd=sqrt(.5*(var.xtreated+var.xcontrol));
  std.diff.before.matching=(mean(xtreated)-mean(xcontrol))/combinedsd;
  nostratum=length(unique(stratum.myindex))-1*max(stratum.myindex==0);
  if(max(stratum.myindex==0)==0){
    stratumlist=sort(unique(stratum.myindex))
  }
  if(max(stratum.myindex==0)==1){
    templist=sort(unique(stratum.myindex))
    stratumlist=templist[-1]
  }
  diff.in.stratum=rep(0,nostratum);
  number.in.stratum=rep(0,nostratum);
  harmonic.weight=rep(0,nostratum)
  for(i in 1:nostratum){
    if(sum(stratum.myindex==stratumlist[i] & treatment==1 & missing==0)==0 | sum(stratum.myindex==stratumlist[i] & treatment==0 & missing==0)==0){
      number.in.stratum[i]=0
    }
    if(sum(stratum.myindex==stratumlist[i] & treatment==1 & missing==0)>0 & sum(stratum.myindex==stratumlist[i] & treatment==0 & missing==0)>0){
      diff.in.stratum[i]=mean(x[stratum.myindex==stratumlist[i] & treatment==1 & missing==0])-mean(x[stratum.myindex==stratumlist[i] & treatment==0 & missing==0]);
      number.in.stratum[i]=sum(stratum.myindex==stratumlist[i])
      harmonic.weight[i]=1/(.5/sum(stratum.myindex==stratumlist[i] & treatment==1)+.5/sum(stratum.myindex==stratumlist[i] & treatment==0))
    }
  }
  std.diff.after.matching=(sum(harmonic.weight*diff.in.stratum)/sum(harmonic.weight))/combinedsd;
  list(std.diff.before.matching=std.diff.before.matching,std.diff.after.matching=std.diff.after.matching);
}

# Also compute balance on logit propensity score
# Xmatmahal$logit.ps=datatemp$logit.ps
# Calculate the standardized differences
std.diff.before=rep(0,ncol(Xmatmahal));
std.diff.after=rep(0,ncol(Xmatmahal));
names(std.diff.before)=names(Xmatmahal[1,]);
names(std.diff.after)=names(Xmatmahal[1,]);
for(i in 1:ncol(Xmatmahal)){
  missing.temp=is.na(Xmatmahal[,i])
  temp.stand.diff=standardized.diff.harmonic.func(Xmatmahal[,i],datatemp$treated,matchedset.index.numeric,missing.temp);
  std.diff.before[i]=temp.stand.diff$std.diff.before.matching;
  std.diff.after[i]=temp.stand.diff$std.diff.after.matching;
}

sd.bf=std.diff.before
sd.af=std.diff.after
options(digits=2)
cbind(sd.bf,sd.af)

covariates=names(sd.bf)

plot.dataframe2=data.frame(stand.diff=c(sd.bf,sd.af),covariates=rep(covariates,2),type=c(rep("Before",length(covariates)),rep("After",length(covariates))))
ggplot(plot.dataframe2,aes(x=stand.diff,y=covariates))+geom_point(size=5,aes(shape=factor(type)))+scale_shape_manual(values=c(4,1))+geom_vline(xintercept=c(-0.2,0.2),lty=2)+
  theme(axis.text.y = element_text(size=6)) 

ggsave(
  filename = "l5.png",
  device = "png", width = 6, height = 4)

#Point estimate for optimal matching and Confidence interval.
reg.formula=update(propscore.model$formula,outcome~treated+matchvec+.)
matched.reg.model=lm(reg.formula,data=datatemp)
summary(matched.reg.model)
# Point estimate of treatment effect
coef(matched.reg.model)[2]
# Confidence interval
confint(matched.reg.model)[2,]



```
The love plot depicts standard differences before and after matching the treatment group of people who did box breathing  vs control group comprising of all people before they started exercises. We performed regression analysis to determine significane  of treatment effect of performing box breathing exercise. We compared used sleep efficacy of both groups as the outcome metric.  The p-value suggests that box breathing improves sleep efficacy, we did a similar analysis for other breathing techiniques using  full matching but did not find any significant treatment effect. 

## Footnotes and Sources

1. Stat 9210 Notes [Matching Code]
2. Brief structured respiration practices enhance mood and reduce physiological arousal (Melis Yilmaz Balban, Eric Neri,
Manuela M. Kogon, ..., Jamie M. Zeitzer, David Spiegel, Andrew D. Huberman) [Boxplots, T-test, Kruskal, Wilcoxon for comparing all the breathworks were generated using the data and code provided in this paper]
3. The American Journal of Medicine [https://www.amjmed.com/]