---
title: "STAT 9210 Final Project"
author: "Pragyat (Prag) Agrawal and Sahil Kulkarni"
date: "2023-04-25"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r Setup, include=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(echo = T, fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output

# Package setup
if(!require("pacman")) install.packages("pacman")

#install.packages("fGarch", repos = "http://cran.us.r-project.org")
#libary("fGarch")

pacman::p_load(plyr, tidyverse, dplyr, ggplot2, ggthemes, data.table, lubridate,
               GGally, RColorBrewer, ggsci, plotROC, usmap,
               plotly, ggpubr, vistime, pander, softImpute, sjmisc, sjPlot, stringr,
               hwwntest, lmtest, spelling, tseries, moments, fGarch, exactRankTests, FSA, psych,
               lme4, lmerTest, nlme, car, readxl)

options(tinytex.verbose = TRUE)


```


# Introduction

Breathing practices are vital for humans as they influence how we feel and shape human behavior. But scientific evidence on breathing and its effect on mind-body connection have been limited. Structured breathing practices have their own share of benefits as they help with enhancing mood and reduce physiological arousal in humans. There have been several studies which have established this. We seek to study this same effect and replicate its result by applying the concepts we learned in class to find if long breathing exercises can help people reach mindfulness and reduce stress and anxiety. 

# Study

We looked into a study by Cell Reports Medicine which according to their website cell.com is "premium open-access journal from Cell Press publishing cutting-edge research in translational and clinical biomedical sciences that inform and influence human health and medicine."


```{r}
filename = read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

view(filename)
head(filename)
```

## Participants

Throughout the study, 140 potential targets were consented but 114 were invited to participate in the study. There was some attrition because of pandemic related reasons or loss of contact with the target. According to the study, The general ease of following instructions and performing the interventions and subjective experience of the interventions were assessed by an optional debriefing questionnaire at the end of the study. 

## Box Study
```{r}
library(readxl)
library(lme4)

library(lmerTest)
library(tidyverse)
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

filename[filename=='.'] <- NA

data_set <- filename[, c(1, 2, 3, 4, 6)]

data_set <- data_set %>% drop_na()

data_set$`Post PANAS Positive Affect` <- as.numeric(data_set$`Post PANAS Positive Affect`)
data_set$`Pre PANAS Positive Affect` <- as.numeric(data_set$`Pre PANAS Positive Affect`)
data_set$`Round 1 Exercise` <- as.factor(data_set$`Round 1 Exercise`)
data_set$`Days from  Round1 Day1` <- as.numeric(data_set$`Days from  Round1 Day1`)


data_set$change <- data_set$`Post PANAS Positive Affect` - data_set$`Pre PANAS Positive Affect`

model = lmer(change ~ `Days from  Round1 Day1` + `Round 1 Exercise` * `Days from  Round1 Day1` +
               `Round 1 Exercise` + (1|ScrSubjectID),
             data=data_set,
             REML=TRUE)
summary(model)

anova(model)
```
```{r}
#Trait anxiety score analysis
library('ggpubr')
library(readxl)
trait <- data.frame(matrix(ncol =5, nrow = 120 ))

#colnames(trait) <- c("subjectID", "Pre Trait", "Post Trait", "Group")

colnames(trait) <- c("subjectID", "Pre Trait", "Post Trait", "Group", "Type")

# filename = BWPilot_CombinedData_20210308
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")


for (i in 1:length(unique(filename$`ScrSubjectID`))){
  print(i)
  id <- unique(filename$`ScrSubjectID`)[i]
  print (id)
  b <- filename$`ScrSubjectID` == as.numeric(id) 
  a <- filename[b,] # data for each individuaal subject
  
  
  round1pre<- subset(a, `Days from  Round1 Day1`<1)
  #ind<- a$`RSP Day` ==28 & a$Round == 1
  # ind<- a$`RSP Day` ==1 & a$Round == 2

  # postind <- which(ind, arr.ind = TRUE, useNames = TRUE)
  # if (length(postind)==0)next()
  # round1post <- a[(postind-5):postind,]
  round1post<- subset(a, `Days from  Round1 Day1`> 1 &`Days from  Round1 Day1`< 31 )
  prepromisind <-round1pre$`Full STAI Trait Anxiety` != "."
  prepromis <- as.numeric(round1pre$`Full STAI Trait Anxiety`[prepromisind])
  postpromisind <-round1post$`Full STAI Trait Anxiety` != "."
  postpromis <- as.numeric(round1post$`Full STAI Trait Anxiety`[postpromisind])
  trait$subjectID [i] <- id
  if (length(prepromis)>0) {
    trait$`Pre Trait` [i] <- prepromis[length(prepromis)]
    
  }
  
  if (length(postpromis)>0) {
    trait$`Post Trait` [i] <- postpromis[length(postpromis)]
    
  }
  trait$Group [i] <-  a$`Round 1 Exercise`[1]
  if (trait$Group[i] == 'Mindful Meditation'){trait$Type[i]<- 'Mindfulness'}
    else {trait$Type[i]<- 'Breathwork'}
  
}
effect_trait <- as.numeric(trait$`Post Trait`) - as.numeric(trait$`Pre Trait`)

trait['delta trait']<- effect_trait


b <- trait$Group == "Mindful Meditation"
trait_medi <- trait[b,]

b <- trait$Group == "Box Breathing"
trait_box <- trait[b,]

b <- trait$Group == "Super Oxygenation"
trait_superox <- trait[b,]

b <- trait$Group == "Slow Breathing"
trait_slow <- trait[b,]
####

effect_superox <- as.numeric(trait_superox$`Post Promis`)- as.numeric(trait_superox$`Pre Promis`)
effect_medi <- as.numeric(trait_medi$`Post Promis`)- as.numeric(trait_medi$`Pre Promis`)
effect_box <- as.numeric(trait_box$`Post Promis`)- as.numeric(trait_box$`Pre Promis`)
effect_slow<- as.numeric(trait_slow$`Post Promis`) - as.numeric(trait_slow$`Pre Promis`)


wilcox.test(trait_superox$`Post Trait`,trait_superox$`Pre Trait`, paired = TRUE)
wilcox.test(trait_slow$`Post Trait`,trait_slow$`Pre Trait`, paired = TRUE)
wilcox.test(trait_box$`Post Trait`,trait_box$`Pre Trait`, paired = TRUE)
wilcox.test(trait_medi$`Post Trait`,trait_medi$`Pre Trait`, paired = TRUE)





#STATS

kruskal.test(`delta trait`~Group, data=trait)
wilcox.test(trait_slow$`Post Trait`,trait_slow$`Pre Trait`, paired = TRUE) #WILCOX IS 2 SAMPLE ONLY, KRUSKAL WALLIS MORE) , ~


##box plot for round and round 2


#round 1
trait <- trait [complete.cases(trait),]
trait$Group<- factor(trait$Group, levels = c('Mindful Meditation', 'Slow Breathing', 'Box Breathing', 'Super Oxygenation'))
e <- ggplot(trait, aes(x = Group, y = `delta trait`))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = "grey20"), size = 2, position = position_jitterdodge(),show.legend = F) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())

## by type t test

traitmed<- subset(trait, Type =='Mindfulness')
traitbreath <- subset(trait, Type == 'Breathwork')
t.test(traitmed$`delta trait`, y =traitbreath$`delta trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)


t.test(traitmed$`Pre Trait`, y =traitmed$`Post Trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.95)

t.test(traitbreath$`Pre Trait`, y =traitbreath$`Post Trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.95)

traitmed<- traitmed[complete.cases(traitmed),]
mean (traitmed$`delta trait`)/sd(traitmed$`Pre Trait`)
traitbreath<- traitbreath[complete.cases(traitbreath),]
mean (traitbreath$`delta trait`)/sd(traitbreath$`Pre Trait`)


#wilcoxontest

wilcox.test(traitbreath$`Post Trait`,traitbreath$`Pre Trait`, paired = TRUE)
wilcox.test(traitbreath$`delta trait`,traitmed$`delta trait`, paired = FALSE)



plot.new() ## this plot doesnt look good
for (i in 1:nrow(traitbreath)){
  if (is.na(traitbreath$`Pre Trait`[i])|is.na(traitbreath$`Post Trait`[i])) next
  par(new=T)
  plot(c(traitbreath$`Pre Trait`[i],traitbreath$`Post Trait` [i]), type= 'o', ylab = 'Trait Anxiety', xlab = 'All Breathwork', xlim=c(0,3),ylim=c(20,80), xaxt='n')
}

axis(1, at=c(1,2), labels=c('Before', 'After'))

trait$Type<- factor(trait$Type, levels = c('Mindfulness', 'Breathwork'))
trait<- trait[complete.cases(trait),]
e <- ggplot(trait, aes(x = Type, y = `delta trait`))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = "grey20"), size = 2, position = position_jitterdodge(),show.legend = F) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())

kruskal.test(`delta trait`~Type, data = trait)


#ttest by group

traitbreath <- subset(trait, Group =='Box Breathing')
traitmed<- subset(trait, Group =='Super Oxygenation')


t.test(traitmed$`delta trait`, y =traitbreath$`delta trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)




```



```{r}
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

model = lmer(Length ~ Hand + (1|Individual),
             data=Data,
             REML=TRUE)

anova(model)
```

