---
title: "STAT 9210 Final Project"
author: "Pragyat (Prag) Agrawal and Sahil Kulkarni"
date: "2023-04-25"
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---


```{r Setup, include=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(echo = T, fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output

# Package setup
if(!require("pacman")) install.packages("pacman")

#install.packages("fGarch", repos = "http://cran.us.r-project.org")
#libary("fGarch")

pacman::p_load(plyr, tidyverse, dplyr, ggplot2, ggthemes, data.table, lubridate,
               GGally, RColorBrewer, ggsci, plotROC, usmap,
               plotly, ggpubr, vistime, pander, softImpute, sjmisc, sjPlot, stringr,
               hwwntest, lmtest, spelling, tseries, moments, fGarch, exactRankTests, FSA, psych,
               lme4, lmerTest, nlme, car, readxl)

options(tinytex.verbose = TRUE)


```


# Introduction

Breathing practices are vital for humans as they influence how we feel and shape human behavior. But scientific evidence on breathing and its effect on mind-body connection have been limited. Structured breathing practices have their own share of benefits as they help with enhancing mood and reduce physiological arousal in humans. There have been several studies which have established this. We seek to study this same effect and replicate its result by applying the concepts we learned in class to find if long breathing exercises can help people reach mindfulness and reduce stress and anxiety. 

# Study

We looked into a study by Cell Reports Medicine which according to their website cell.com is "premium open-access journal from Cell Press publishing cutting-edge research in translational and clinical biomedical sciences that inform and influence human health and medicine."


```{r}
filename = read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

view(filename)
head(filename)
```

## Participants

Throughout the study, 140 potential targets were consented but 114 were invited to participate in the study. There was some attrition because of pandemic related reasons or loss of contact with the target. According to the study, The general ease of following instructions and performing the interventions and subjective experience of the interventions were assessed by an optional debriefing questionnaire at the end of the study. 

## Treatment

The experiment put the 114 participants to a breathing and meditation protocol and measured the affect of that on their state anxiety. Specifically, it measured positive affect (Positive And Negative Affect Schedule (PANAS)) and state anxiety (State-Trait Anxiety Inventory (STAI)) scores on each participant before and after the experiment. Mindfulness meditation and breathwork groups both experienced an increase in daily positive affect.


## Data Visualization

```{r}
attach(filename)
filename$`Sleep Score` <- as.numeric(filename$`Sleep Score`)
filename$`Hours of Sleep` <- as.numeric(filename$`Hours of Sleep`)
# create new variables with different names
my_hours <- filename$`Hours of Sleep`
my_score <- filename$`Sleep Score`

# use the original variables with the `::` operator
ggplot(filename, aes(x = my_hours, y = my_score)) + 
  geom_point() + 
  geom_smooth(method = lm, color = 'red', se = F) + 
  ylim(NA, 100) + 
  annotate("text",x=10,y=10,label=(paste0("slope==",coef(lm(filename$`Sleep Score`~filename$`Hours of Sleep`))[2])),parse=TRUE)
```
This graphs shows that sleep score is positively related to the number of hours a person sleeps. This means a higher number of hours sleeping, corresponds to a higher sleep score. This makes sense as the quality of sleep increases with the quantity of sleep a person gets. 


```{r}
colnames(filename)
```

```{r}
library(ggplot2)
library(tidyr)
library(cowplot)

variables <- filename[,4:9]
variables <- apply(variables, 2, function(x) as.numeric(as.character(x)))
variables <- variables[!apply(is.na(variables) | variables == "", 1, all),]
variables_df <- as.data.frame(variables)

group1 <- variables_df[, c(1, 2, 5)]
group2 <- variables_df[, c(3, 4, 6)]

group1_melted <- gather(group1, key = "variable", value = "value")
group2_melted <- gather(group2, key = "variable", value = "value")

p1 <- ggplot(group1_melted, aes(x = variable, y = value)) +
  geom_boxplot(position = position_dodge(0.75)) +
  xlab("") +
  ylab("Group 1") + 
  theme(axis.text.x = element_text(angle = 10))


p2 <- ggplot(group2_melted, aes(x = variable, y = value)) +
  geom_boxplot(position = position_dodge(0.75)) +
  xlab("") +
  ylab("Group 2")+ 
  theme(axis.text.x = element_text(angle = 10))

plot_grid(p1, p2, ncol = 2)

```
Looking at the results, we see the breathwork group had a notably higher increase in daily positive affect, a reduction in daily negative affect and a reduction in the state anxiety. All of these indicate that there is some positive effect because of the breathing and meditation protocol that has been assigned to the participants. This is proven by the regression below which shows Mindful Meditation as a significant indicator for a reduction in anxiety and an increase in positive effect. 

## Box Study

```{r}
library(readxl)
library(lme4)

library(lmerTest)
library(tidyverse)
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

filename[filename=='.'] <- NA

data_set <- filename[, c(1, 2, 3, 4, 6)]

data_set <- data_set %>% drop_na()

data_set$`Post PANAS Positive Affect` <- as.numeric(data_set$`Post PANAS Positive Affect`)
data_set$`Pre PANAS Positive Affect` <- as.numeric(data_set$`Pre PANAS Positive Affect`)
data_set$`Round 1 Exercise` <- as.factor(data_set$`Round 1 Exercise`)
data_set$`Days from  Round1 Day1` <- as.numeric(data_set$`Days from  Round1 Day1`)


data_set$change <- data_set$`Post PANAS Positive Affect` - data_set$`Pre PANAS Positive Affect`

model = lmer(change ~ `Days from  Round1 Day1` + `Round 1 Exercise` * `Days from  Round1 Day1` +
               `Round 1 Exercise` + (1|ScrSubjectID),
             data=data_set,
             REML=TRUE)
summary(model)

anova(model)
```
```{r Trait_z_1, include=FALSE}
#Trait anxiety score analysis
library('ggpubr')
library(readxl)
trait <- data.frame(matrix(ncol =5, nrow = 120 ))

#colnames(trait) <- c("subjectID", "Pre Trait", "Post Trait", "Group")

colnames(trait) <- c("subjectID", "Pre Trait", "Post Trait", "Group", "Type")

# filename = BWPilot_CombinedData_20210308
filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")


for (i in 1:length(unique(filename$`ScrSubjectID`))){
  print(i)
  id <- unique(filename$`ScrSubjectID`)[i]
  print (id)
  b <- filename$`ScrSubjectID` == as.numeric(id) 
  a <- filename[b,] # data for each individuaal subject
  
  
  round1pre<- subset(a, `Days from  Round1 Day1`<1)
  #ind<- a$`RSP Day` ==28 & a$Round == 1
  # ind<- a$`RSP Day` ==1 & a$Round == 2

  # postind <- which(ind, arr.ind = TRUE, useNames = TRUE)
  # if (length(postind)==0)next()
  # round1post <- a[(postind-5):postind,]
  round1post<- subset(a, `Days from  Round1 Day1`> 1 &`Days from  Round1 Day1`< 31 )
  prepromisind <-round1pre$`Full STAI Trait Anxiety` != "."
  prepromis <- as.numeric(round1pre$`Full STAI Trait Anxiety`[prepromisind])
  postpromisind <-round1post$`Full STAI Trait Anxiety` != "."
  postpromis <- as.numeric(round1post$`Full STAI Trait Anxiety`[postpromisind])
  trait$subjectID [i] <- id
  if (length(prepromis)>0) {
    trait$`Pre Trait` [i] <- prepromis[length(prepromis)]
    
  }
  
  if (length(postpromis)>0) {
    trait$`Post Trait` [i] <- postpromis[length(postpromis)]
    
  }
  trait$Group [i] <-  a$`Round 1 Exercise`[1]
  if (trait$Group[i] == 'Mindful Meditation'){trait$Type[i]<- 'Mindfulness'}
    else {trait$Type[i]<- 'Breathwork'}
  
}
```

```{r Trait_z_2}

effect_trait <- as.numeric(trait$`Post Trait`) - as.numeric(trait$`Pre Trait`)

trait['delta trait']<- effect_trait


b <- trait$Group == "Mindful Meditation"
trait_medi <- trait[b,]

b <- trait$Group == "Box Breathing"
trait_box <- trait[b,]

b <- trait$Group == "Super Oxygenation"
trait_superox <- trait[b,]

b <- trait$Group == "Slow Breathing"
trait_slow <- trait[b,]
####

effect_superox <- as.numeric(trait_superox$`Post Promis`)- as.numeric(trait_superox$`Pre Promis`)
effect_medi <- as.numeric(trait_medi$`Post Promis`)- as.numeric(trait_medi$`Pre Promis`)
effect_box <- as.numeric(trait_box$`Post Promis`)- as.numeric(trait_box$`Pre Promis`)
effect_slow<- as.numeric(trait_slow$`Post Promis`) - as.numeric(trait_slow$`Pre Promis`)


wilcox.test(trait_superox$`Post Trait`,trait_superox$`Pre Trait`, paired = TRUE)
wilcox.test(trait_slow$`Post Trait`,trait_slow$`Pre Trait`, paired = TRUE)
wilcox.test(trait_box$`Post Trait`,trait_box$`Pre Trait`, paired = TRUE)
wilcox.test(trait_medi$`Post Trait`,trait_medi$`Pre Trait`, paired = TRUE)





#STATS

kruskal.test(`delta trait`~Group, data=trait)
wilcox.test(trait_slow$`Post Trait`,trait_slow$`Pre Trait`, paired = TRUE) #WILCOX IS 2 SAMPLE ONLY, KRUSKAL WALLIS MORE) , ~


##box plot for round and round 2


#round 1
trait <- trait [complete.cases(trait),]
trait$Group<- factor(trait$Group, levels = c('Mindful Meditation', 'Slow Breathing', 'Box Breathing', 'Super Oxygenation'))
e <- ggplot(trait, aes(x = Group, y = `delta trait`))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = "grey20"), size = 2, position = position_jitterdodge(),show.legend = F) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())

## by type t test

traitmed<- subset(trait, Type =='Mindfulness')
traitbreath <- subset(trait, Type == 'Breathwork')
t.test(traitmed$`delta trait`, y =traitbreath$`delta trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)


t.test(traitmed$`Pre Trait`, y =traitmed$`Post Trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.95)

t.test(traitbreath$`Pre Trait`, y =traitbreath$`Post Trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = TRUE, var.equal = FALSE,
       conf.level = 0.95)

traitmed<- traitmed[complete.cases(traitmed),]
mean (traitmed$`delta trait`)/sd(traitmed$`Pre Trait`)
traitbreath<- traitbreath[complete.cases(traitbreath),]
mean (traitbreath$`delta trait`)/sd(traitbreath$`Pre Trait`)


#wilcoxontest

wilcox.test(traitbreath$`Post Trait`,traitbreath$`Pre Trait`, paired = TRUE)
wilcox.test(traitbreath$`delta trait`,traitmed$`delta trait`, paired = FALSE)



plot.new() ## this plot doesnt look good
for (i in 1:nrow(traitbreath)){
  if (is.na(traitbreath$`Pre Trait`[i])|is.na(traitbreath$`Post Trait`[i])) next
  par(new=T)
  plot(c(traitbreath$`Pre Trait`[i],traitbreath$`Post Trait` [i]), type= 'o', ylab = 'Trait Anxiety', xlab = 'All Breathwork', xlim=c(0,3),ylim=c(20,80), xaxt='n')
}

axis(1, at=c(1,2), labels=c('Before', 'After'))

trait$Type<- factor(trait$Type, levels = c('Mindfulness', 'Breathwork'))
trait<- trait[complete.cases(trait),]
e <- ggplot(trait, aes(x = Type, y = `delta trait`))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = "grey20"), size = 2, position = position_jitterdodge(),show.legend = F) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())

kruskal.test(`delta trait`~Type, data = trait)


#ttest by group

traitbreath <- subset(trait, Group =='Box Breathing')
traitmed<- subset(trait, Group =='Super Oxygenation')


t.test(traitmed$`delta trait`, y =traitbreath$`delta trait`,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)




```
We have conducted Wilcoxin and Kruskal-Wallis rank sum test, whose output is above. The wilcoxin test for the super-oxygenation group is significant at the 0.10 level. The Kruskal-Wallis test is not significant which may indicate that there is no evidence that the medians of the pre and post groups are significantly different. However, it does not say that there are no differences in the groups. It may be the case that the differences are too small to be detected.

The paired t-test for pre and post traitbreath shows that the difference in factor is significant. 

This graphs shows the before and after trait anxiety. This graph doesn't show this clearly, but it seems that there is a reduction in before and after state anxiety. We have to do further analysis to prove this.






```{r SlopesRR_1, include=FALSE}
library("ggpubr")
library(readxl)

filename <- read_excel("BWPilot_CombinedData_20210803_fordryad_addvars_cleaned_noround2_v3.xlsx")

slopes<- data.frame(matrix (ncol = 11, nrow = 120))

for (i in 1:length(unique(filename$`ScrSubjectID`))){
  print(i)
  id <- unique(filename$`ScrSubjectID`)[i]
  print (id)
  b <- filename$`ScrSubjectID` == as.numeric(id) 
  a <- filename[b,] # data for each individuaal subject
  c<- a #round 1
  
  c[c=='.'] <- NA
  
  if (nrow(c) == 0){
    next()
  }
  
  if (sum(is.na(c$RHR))==length(c$RHR))
    next()
  
  #  if (sum(!is.na(c$RHR))<2) 
   #   next()
   
  if (sum(!is.na(c$`Pre PANAS Positive Affect`))<2) 
      next()
  
  z<- c[,c(2,4,5,6,7,8,9,15,16,20)] # extract hrv, rhr, respiratory rate, sleep scores
  z<- unique.data.frame(z)

  #reg1 <- lm(c$HRV ~ c$`Days from  Round1 Day1`,data=c, na.action = na.omit)
  slope <- lm(formula = z$HRV ~ z$`Days from  Round1 Day1`)
  slopeHRV<-slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  slope <- lm(formula = z$RHR ~ z$`Days from  Round1 Day1`)
  slopeRHR<- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  slope <- lm(formula = z$`Respiration Rate` ~ z$`Days from  Round1 Day1`)
  slopeRR <- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  # slope <- lm(formula = z$`Sleep Score` ~ z$`Days from  Round1 Day1`)
  # slopesleepscore <- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  # slope <- lm(formula = z$`Hours of Sleep` ~ z$`Days from  Round1 Day1`)
  # slopehoursofsleep <- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  # slope <- lm(formula = z$`Sleep Efficiency` ~ z$`Days from  Round1 Day1`)
  # slopesleepefficiency <- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  # #slope <- lm(formula = z$`Pre STAI State Anxiety` ~ z$`Days from  Round1 Day1`)
  # #slopeprestate <- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
 slope <- lm(formula = c$`Pre PANAS Positive Affect`~ z$`Days from  Round1 Day1`)
slopeprepanaspos<- slope[['coefficients']][['z$`Days from  Round1 Day1`']]
  
  
  z$deltapanaspos<- as.numeric(z$`Post PANAS Positive Affect`)- as.numeric(z$`Pre PANAS Positive Affect`)
    slope <- lm(formula = z$deltapanaspos~ as.numeric(z$`Days from  Round1 Day1`))
    slopedeltapanaspos <- slope[['coefficients']][2]
  # 
  
  
  #plot(z$`Days from  Round1 Day1`, z$RR, main = id, xlab = "Days from Round 1", ylab = 'RHR' , pch = 20, cex = 4)
#  abline(lm(z$RHR ~ z$`Days from  Round1 Day1`))

 # cor.test(as.numeric(z$`Days from  Round1 Day1`), as.numeric(z$RHR), method="pearson")
  
  
  
 # plot(z$`Days from  Round1 Day1`, z$`Respiration Rate`, main = id, xlab = "Days from Round 1", ylab = 'Respiration Rate' ,pch = 20, cex = 4, bty = 'n')
#  abline(lm(z$`Respiration Rate` ~ z$`Days from  Round1 Day1`))
 # cor.test(as.numeric(z$`Days from  Round1 Day1`), as.numeric(z$`Respiration Rate`), method="pearson")
  
  #reg1 <- lm(z$RHR ~ z$`Days from  Round1 Day1`,data=z) 
  
  slopes [i,1] <- id
  slopes [i,2] <- slopeHRV
  slopes [i,3] <- slopeRHR
  slopes [i,4] <- slopeRR
 slopes [i,5] <- slopedeltapanaspos
  slopes [i,6] <- c$`Round 1 Exercise`[1]
 # slopes [i,7] <- c$Age[1]
 # slopes [i,8] <- c$Gender[1]
 # slopes [i,7] <- slopesleepscore
#  slopes [i,8] <- slopesleepefficiency
 # slopes [i,9] <- slopehoursofsleep
#  slopes[i,10]<-slopeprestate
  slopes [i,11]<- slopeprepanaspos
  
  colnames(slopes)<- c("subjectID", "slopeHRV", "slopeRHR", "slopeRR","slopedeltapanaspos","Group", "SleepScore", "SleepEfficiency", "HoursOfSleep", "slopeprestate", "slopeprepanaspos")
  
}

```

```{r SlopesRR_2}

#remove any empty columns

slopes <- slopes[,colSums(is.na(slopes))<nrow(slopes)]
slopes_rhrhrvRR_excludinglessthan8datapoints<- slopes # we are not excluding data. 

#statistical tests
#non-parametric

slopes_rhrhrvRR_excludinglessthan8datapoints<- slopes_rhrhrvRR_excludinglessthan8datapoints[complete.cases(slopes_rhrhrvRR_excludinglessthan8datapoints),]
kruskal.test(slopeRR~Group, data = slopes_rhrhrvRR_excludinglessthan8datapoints)
pairwise.wilcox.test(slopes_rhrhrvRR_excludinglessthan8datapoints$slopeRR, slopes_rhrhrvRR_excludinglessthan8datapoints$Group, paired = FALSE)

    
  for (i in 1:nrow(slopes_rhrhrvRR_excludinglessthan8datapoints)){
  if (slopes_rhrhrvRR_excludinglessthan8datapoints$Group[i] =="Mindful Meditation"){ slopes_rhrhrvRR_excludinglessthan8datapoints$Type[i]<- 'Mindfulness'}
  # else if (c$`Round 1 Exercise` == 'Box Breathing'){ weekly$Type[i]<- 'Box Breathing'}
  else {slopes_rhrhrvRR_excludinglessthan8datapoints$Type[i]<- 'Breathing'}
  }
  slopes_rhrhrvRR_excludinglessthan8datapoints$Type<- factor(slopes_rhrhrvRR_excludinglessthan8datapoints$Type, levels = c("Mindfulness", "Breathing"))
  
  kruskal.test(slopeRR~Type, data = slopes_rhrhrvRR_excludinglessthan8datapoints)
  
  rrm <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Mindfulness")
  rrb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Breathing")
  rrsb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Group == "Slow Breathing")
  rrbb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Group == "Box Breathing")
  rrso <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Group == "Super Oxygenation")
  
  deltapanasposm <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Mindfulness")
  deltapanasposb <- subset(slopes_rhrhrvRR_excludinglessthan8datapoints, Type == "Breathing")
  
# t test
  
  t.test(rrm$slopeRR, y =rrb$slopeRR,
         alternative = c ("two.sided", "less", "greater"),
         mu = 0, paired = FALSE, var.equal = FALSE,
         conf.level = 0.95)
  signal<- slopes_rhrhrvRR_excludinglessthan8datapoints
  signal <- signal [complete.cases(signal),]
  signal$Type<- factor(signal$Type, levels = c('Mindfulness', 'Breathing'))
  signal$Group<- factor(signal$Group, levels = c('Mindful Meditation', 'Slow Breathing', 'Box Breathing', 'Super Oxygenation'))
  print(i)
  e <- ggplot(signal, aes(x = Type, y = slopeRR))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = 'red'), size = 2, position = position_jitterdodge(),show.legend = F) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
  e
  e <- ggplot(signal, aes(x = Group, y = slopeRR))+geom_boxplot(alpha = 0.3)
  e
  # t.test(rrb$HoursOfSleep)
  # t.test(rrb$SleepEfficiency)
  
  # t.test(deltapanasposm$slopePANASPos, y =deltapanasposb$slopePANASPos,
  #        alternative = c("two.sided", "less", "greater"),
  #        mu = 0, paired = FALSE, var.equal = FALSE,
  #        conf.level = 0.95)
  # 
  # pairwise.wilcox.test(slopes_rhrhrvRR_excludinglessthan8datapoints$slopePANASPos, slopes_rhrhrvRR_excludinglessthan8datapoints$Type, paired = FALSE)
  e <- ggplot(slopes_rhrhrvRR_excludinglessthan8datapoints, aes(x = Group, y = slopeHRV))+geom_boxplot(alpha = 0.3)+geom_point (aes(colour = Group), size = 2, position = position_jitterdodge(),show.legend = T) + theme(axis.text.y = element_text(color = "grey20", size = 30, angle = 0, hjust = .5, vjust = .5, face = "plain"),  axis.text.x = element_text(color = "grey20", size =20, angle = 0, hjust = .5, vjust = .5, face = "plain"),axis.title = element_blank(), axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
  e
  # ## add in dayscompleted
  

  # for (i in 1:nrow(slopes_rhrhrvRR_excludinglessthan8datapoints)){
  #   id<- slopes_rhrhrvRR_excludinglessthan8datapoints$subjectID[i]
  # dayscompleted <- dayscompleted_cumulative[dayscompleted_cumulative$subjectID==id,3]
  # slopes_rhrhrvRR_excludinglessthan8datapoints$dayscompleted[i]<- dayscompleted
  # }
  # ## calculate mean for slow breathing (fig 4
  # write_csv(slopes, 'slopes_rhrhrvRR_excludinglessthan8datapoints.csv')
  # 
```
